\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-04-11 archive.tex
moved expertise content into main paper

\subsection{Programs that manipulate sequences}\label{sequences}
We first apply \system to two classic benchmark domains: list
processing and text editing. In both cases we solve tasks specified by
a input/output examples, starting with a generic functional programming
basis: \code{foldr}, \code{unfold}, \code{if}, \code{map},
\code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0},
\code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and
\code{is-nil}.

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ecc},
each with 15 input/output examples.  In solving these tasks, the
system composed 16 new DSL components, and discovered multiple
higher-order functions. Each round of memory consolidation built on
components discovered in earlier sleep cycles --- for example the
agent first learns the higher-order function \code{filter}, uses
\code{filter} to learn to take the maximum element of a list, then
uses that routine to learn a new component for extracting the
$n^{\text{th}}$ largest element of a list, which it finally uses to
solve a task involving sorting a list of numbers (Figure~\ref{initialExampleDSL}).
This incremental, modular learning of deep hierarchies of DSL
components occurs because of the alternation between code writing
(during waking) and code refactoring (during the consolidation phase
of sleep).

@@

@@

\subsection{Acquiring Domain Expertise}

One interpretation of our system is as a model of the acquisition of
domain expertise. Humans can acquire expertise across many domains --
cooking, coding, music, architecture, painting, tennis, or calculus,
to name a handful of examples, and every child develops expertise in
natural language, intuitive physics, motor control, kinship
relationships, and more. Domain experts learn domain-specific
abstractions, similar to a DSL: for example, a expert chef knows what
combinations of seasonings go together, or an expert mathematician
knows a wide set of useful theorems and lemmas.  Jointly, experts
learn how to recognize when to use these abstractions to solve
problems.  Becoming an expert involves a learning trajectory that
unfolds over relatively long time scales, but has modest data
requirements relative to the dominant machine learning paradigms. For
example, basic competency in cooking or coding might require on the
order of learning a hundred recipes or solving a hundred programming
exercises.  It is this learning regime that we have targeted with
\systemEnding: Learning about a domain from at most several hundred
tasks, but where the learning unfolds over many wake/sleep cycles.

@@

\subsection{Neural Initial Problem Representations}

How do \systemEnding `s representations of problems themselves change, as the model acquires expertise? Within cognitive science, a rich body of evidence suggests that human experts learn to `see` problems differently in their domains -- even before solving a problem, human novices and experts seem to construct fundamentally different initial representations \cite{chi1981categorization}. Novices notice a problem’s `surface features`, the literal objects and details made obvious by the task description. Experts, on the other hand, somehow see past these details to pick out each problem’s `deep structures`, the underlying principles and concepts that govern its ultimate solution \cite{chi2012seeing,chi2006two}.

\systemEnding`s neural recognition model can be seen as forming analogous `initial` problem representations, implicitly encoded within the activations of the neural network. Intuitively, these initial neural representations map problem features to a `gist` over the model’s current domain-specific concepts -- they guide and constrain how the model searches for problem solutions, but are not symbolically structured solutions themselves. To probe how these representations change as the model gains a richer conceptual vocabulary within each domain, we compare how they cluster problems within that domain over the course of \systemEnding`s learning trajectory. Figure \ref{allTSNE} depicts TSNE visualizations of these neural problem representations at the first and last iteration of the algorithm. We color-code each problem using semantic, human-interpretable categories determined apriori by the task designers, but critically, the model itself never has access to these categories. Rather, as \systemEnding acquires domain expertise, the neural network adapts how it transforms each task`s surface features, restructuring its initial problem representation space to draw together problems that share deeper, more abstract conceptual similarities.

@@

>>>>

\subsection{Programs that manipulate sequences}\label{sequences}
We first apply \system to two classic benchmark domains: list
processing and text editing. In both cases we solve tasks specified by
a input/output examples, starting with a generic functional programming
basis, including routines like \code{map}, \code{fold}, \code{cons}, \code{car}, \code{cdr}, etc.
%% : \code{foldr}, \code{unfold}, \code{if}, \code{map},
%% \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0},
%% \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and
%% \code{is-nil}.

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ecc},
each with 15 input/output examples.  In solving these tasks, the
system typically composes around 20 new DSL components, and discovered multiple
higher-order functions. Each round of memory consolidation built on
components discovered in earlier sleep cycles --- for example the
agent first learns the higher-order function \code{filter}, uses
\code{filter} to learn to take the maximum element of a list, then
uses that routine to learn a new component for extracting the
$n^{\text{th}}$ largest element of a list, which it finally uses to
solve a task involving sorting a list of numbers (Figure~\ref{initialExampleDSL}).
This incremental, modular learning of deep hierarchies of DSL
components occurs because of the alternation between code writing
(during waking) and code refactoring (during the consolidation phase
of sleep).

@@

How do \systemEnding's representations of these problems themselves change, as the model acquires expertise? Within cognitive science, a rich body of evidence suggests that human experts learn to `see' problems differently in their domains -- even before solving a problem, human novices and experts seem to construct fundamentally different initial representations \cite{chi1981categorization}. Novices notice a problem’s `surface features', the literal objects and details made obvious by the task description. Experts, on the other hand, somehow see past these details to pick out each problem’s `deep structures', the underlying principles and concepts that govern its ultimate solution \cite{chi2012seeing,chi2006two}.

\systemEnding's neural recognition model can be seen as forming analogous `initial' problem representations, implicitly encoded within the activations of the neural network. Intuitively, these initial neural representations map problem features to a `gist' over the model’s current domain-specific concepts -- they guide and constrain how the model searches for problem solutions, but are not symbolically structured solutions themselves. To probe how these representations change as the model gains a richer conceptual vocabulary within each domain, we compare how they cluster problems within that domain over the course of \systemEnding's learning trajectory. Figure \ref{textExpertiseCluster} depicts TSNE visualizations of these neural problem representations at the first and last iteration of the algorithm. We color-code each problem using semantic, human-interpretable categories determined apriori by the task designers, but critically, the model itself never has access to these categories. Rather, as \system acquires domain expertise, the neural network adapts how it transforms each task's surface features, restructuring its initial problem representation space to draw together problems that share deeper, more abstract conceptual similarities.

\begin{figure}
  \includegraphics[width = \textwidth]{figures/text_tsne.png}
  \caption{Left: TSNE embedding of model's `at-a-glance' (before searching for solution) organization of text tasks (neural net activations), after first wake/sleep cycle. Middle: at-a-glance organization after 20 wake/sleep cycles. Right: TSNE visualization of programs actually used to solve tasks; intuitively, how model organizes tasks after searching for a solution. More examples in Fig.~\ref{allTSNE}}\label{textExpertiseCluster}
\end{figure}

@@

@@

@@

\subsection{Neural Problem Representations}
